---
title: "v4 Annotation for co-expression analysis"
output: html_notebook
---

# In the chuck below, we load in all the essential libraries for the data analysis. 
```{r Load Libraries, echo=FALSE, results=FALSE, include=FALSE}
library(dplyr)
library(dbplyr)
library(Seurat)
library(SeuratDisk)
library(SeuratData)
library(patchwork) # Also being used for Cellchat (but main purpose was something else)
library(hdf5r)
library(tidyverse)
library(gsubfn)
library(ggplot2)
library(RColorBrewer)
library(writexl)

# For annotation, gene name conversions
library(rPanglaoDB)
library(AnnotationHub)
library(ensembldb)
library(clusterProfiler)
library(org.Hs.eg.db)
library(org.Ss.eg.db)
library(DropletUtils)
library(biomaRt)

# to download external library
library(devtools)

# for Monocle3
library(SeuratWrappers)
library(monocle3)

# plotting and data science packages
library(tidyverse)
library(cowplot)
library(patchwork) # also used for cellchat 

# co-expression network analysis packages:
# NOTE: DOWNLAOD THE "FORTRAN compiler" on Mac systems using ARM64. This compiler is not integrated in Xcode.
library(qlcMatrix)  
library(impute) 
library(preprocessCore) 
library(WGCNA)
library(hdWGCNA)

# deconvolution
library(Biobase) # install this package if needed (for simulated data)
library(BisqueRNA)

# Cellchat and dependencies:
library(NMF)
library(circlize)
library(ComplexHeatmap)
library(presto)
library(CellChat)
```


#### Seurat v4 for WGCNA comparability ####

#### Porcine Ileum Tissue (Wiarda) ####

# Opening the files
```{r, opening the Seurat object, echo=FALSE}
parent_directory = "/Users/hamid/Desktop/ABG MSc project/Raw data/Ileum tissue/Wiarda - Ileum Tissue (non-immune)"
project_name = "Ileum Tissue (WIARDA et al., 2023)"

sobj <- LoadH5Seurat("/Users/hamid/Desktop/ABG MSc project/Raw data/Ileum tissue/Wiarda - Ileum Tissue (non-immune)/Ileum_NonImmuneOnly.h5seurat") # open this from Wiarda script or put the count matrix in the directory of the current script
raw <- as.matrix(sobj@assays$RNA@counts)
rm(sobj)
sobj <- CreateSeuratObject(counts = raw, assay = "RNA", project = "ileum_Tiss", min.cells = 0, min.features = 0)
rm(raw)

# UPDATE GENE NAMES
# Connect to the sScrofa database using mart
ensembl <- useMart("ensembl", dataset="sscrofa_gene_ensembl") 
# For all of our ensembl_gene_id, find all gene symbols
bm <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), values=rownames(sobj), mart=ensembl)
# Find all the ENSEMBL ids in our Seurat object
ENS <- sobj@assays$RNA@counts@Dimnames[[1]][c(grep("ENS", rownames(sobj)))]
# Find the positions of our Seurat ENSEMBL ids in the bm output
ENSEMBL_IDs.pos <- na.omit(c(match(c(paste(ENS)), bm[,1]))) # and remove NAs (ENS with no gene IDs)
# Now find the GeneIDs in bm using the ENSEMBL_IDs.pos
geneIDs <- bm[,2][ENSEMBL_IDs.pos] 
# Since not all ENSEMBL_IDs have true gene name, we need to find the index of geneIDs with no gene name and remove their corresponding ENSEMBL_IDs
# To do this, we will remove genes with no names, and then find the index of these genesID in bm in order to find the corresponding ENSEMBL ids
geneIDs <- as.character(geneIDs[geneIDs !=""])
# Ensure that we don't replace ENSEMBL ids with gene names that could result in duplicates
duplicates <- generics::intersect(geneIDs, rownames(sobj)) # find duplicated between new gene-ids with existing ones in the Seurat object
duplicates.pos <- c(match(c(paste(duplicates)), geneIDs)) # Find postion of the duplicate
geneIDs <- geneIDs[-duplicates.pos] # and remove the duplicarte

geneIDs.pos <- c(match(c(paste(geneIDs)), bm[,2])) # find position of gene IDs in them
ENSEMBL_IDs <- bm[,1][geneIDs.pos] # find the ENSEMBL_IDs using the positions of the gene IDs
# Now we have found all the ENSEMBL ids that have a annotated gene name (winch are also found in our Seurat object)
# We will their final positions in our Seurat object and finally replace them with the true gene ID
# NOTE: The order of the genes are conserved using the method above
final.pos <- c(match(c(paste(ENSEMBL_IDs)), rownames(sobj)))

# replace the name of that genes for all slots
sobj@assays$RNA@counts@Dimnames[[1]][c(final.pos)] <- c(paste(geneIDs))
sobj@assays$RNA@data@Dimnames[[1]][c(final.pos)] <- c(paste(geneIDs))
rownames(sobj@assays$RNA@meta.features)[c(final.pos)] <- c(paste(geneIDs))

# Change the name of the ENSSSCG00000026302 to MKI67
MKI67_gene <- "ENSSSCG00000026302"
# position/index of all genes for all slots
genes <- sobj@assays$RNA@counts@Dimnames[[1]]
# find position of ENSSSCG00000026302
MKI67.position.sobj <- c(match(c(paste(MKI67_gene)), genes))
# replace the name of that gene to MUC2 fro all slots
sobj@assays$RNA@counts@Dimnames[[1]][[MKI67.position.sobj]] <- c(paste("MKI67")) 
sobj@assays$RNA@data@Dimnames[[1]][[MKI67.position.sobj]] <- c(paste("MKI67")) 
rownames(sobj@assays$RNA@meta.features)[MKI67.position.sobj] <- c(paste("MKI67"))
```

#### Standard Workflow ####

# Wiarda et all have already gone throught the standar workflow, thus we will not do it again as their Seaurat object contains all metadata slots.
# Since, we want to apply our own workflow, in the chuck below we will apply the our previous workflow starting with the normalization, feauture selection, sclaing, PCA, clustering and DGE analysis
# For the clustering, we will analyze all clusters manually and select the best.
```{r applying workflow on filtered data, fig.height=4, fig.width=5, echo=FALSE}
# Normalization
sobj <- NormalizeData(sobj, normalization.method = "LogNormalize", scale.factor = 10000) 
# Most variable feature selection
sobj <- FindVariableFeatures(sobj, selection.method = "vst", nfeatures = 2300) # 2500 genes seems suitable to find the essential clusters in ileum tissue
# Scale the data
all.genes.sobj <- rownames(sobj)
sobj <- ScaleData(sobj, features = all.genes.sobj) 
# fix the gene names in scaled data.slot
rownames(sobj@assays[["RNA"]]@scale.data)[MKI67.position.sobj] <- c(paste("MKI67"))
rownames(sobj@assays[["RNA"]]@scale.data)[c(final.pos)] <- c(paste(geneIDs))
# PCA using selected features
sobj <- RunPCA(sobj, features = VariableFeatures(object = sobj)) 
pc.stdev.percentage <- sobj[["pca"]]@stdev / sum(sobj[["pca"]]@stdev) * 100 
cumu.percentage <- cumsum(pc.stdev.percentage) 
co1 <- which(cumu.percentage > 90 & pc.stdev.percentage < 5)[1] 
co2 <- sort(which((pc.stdev.percentage[1:length(pc.stdev.percentage) - 1] - pc.stdev.percentage[2:length(pc.stdev.percentage)]) > 0.05), decreasing = T)[1]  
pcs <- min(co1, co2) 
pc.dims <- 1:pcs
# UMAP/T-SNE clustering
sobj <- FindNeighbors(sobj, dims = pc.dims)
sobj <- FindClusters(sobj, resolution = c(1.0)) # Res 1.0 is final decision for Ileum Tissue 
sobj <- RunUMAP(sobj, dims = pc.dims) # UMAP
```


# In the chuck below, we choose the preleminary cluster based on expectation and visualize
```{r final cluster, fig.height=4, fig.width=5, echo=FALSE} 
# setting final cluster
Idents(sobj) <- "seurat_clusters" #1 stock
```


At this point, we have classified our porcine ileum tissue scRNA-seq dataset and been able to classify clusters into the following cell lineages:

* Endothelium                          (cluster 4)
* Mesenchyme                           (cluster 6)
* Prolific Stem/Prog                   (cluster 3)
* EC Prog                              (cluster 5) 
* PC                                   (cluster 9) 
* GC                                   (Cluster 1)
* Developing EC                        (cluster 0, 10)
* Early EC                             (cluster 7)
* Late EC                              (cluster 2)
* EEC                                  (Cluster 11, 12)
* TC                                   (Cluster 8)


#### Annotate/Heatmap/Dotplot of DGEs per cluster/cell type ####


```{r, annotation plot, fig.height=4, fig.width=4.3, echo=FALSE}
bcs <- as.data.frame(colnames(sobj))
colnames(bcs) <- 'barcode'
bcs$cellID <- rep('_', nrow(bcs))
df <- as.data.frame(sobj$seurat_clusters)
# Cluster-based cell annotations not provided in the annotation .txt files:
C4 <- rownames(subset(df, sobj$seurat_clusters %in% c("4")))
C6 <- rownames(subset(df, sobj$seurat_clusters %in% c("6")))
C3 <- rownames(subset(df, sobj$seurat_clusters %in% c("3")))
C5 <- rownames(subset(df, sobj$seurat_clusters %in% c("5")))
C9 <- rownames(subset(df, sobj$seurat_clusters %in% c("9")))
C1 <- rownames(subset(df, sobj$seurat_clusters %in% c("1")))
C0.10 <- rownames(subset(df, sobj$seurat_clusters %in% c("0", "10")))
C7 <- rownames(subset(df, sobj$seurat_clusters %in% c("7")))
C2 <- rownames(subset(df, sobj$seurat_clusters %in% c("2")))
C11.12 <- rownames(subset(df, sobj$seurat_clusters %in% c("11", "12")))
C8 <- rownames(subset(df, sobj$seurat_clusters %in% c("8")))
# Name the cells in the cluster
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C4, 'Endothelium *'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C6, 'Mesenchyme *'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C3, 'Prolific Stem/Prog *'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C5, 'EC Prog *'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C9, 'PC *'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C1, 'GC *'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C0.10, 'Developing EC *'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C7, 'Early EC *'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C2, 'Late EC *'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C11.12, 'EEC *'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C8, 'TC *'))
# update the metadatas slot
rownames(bcs) <- bcs$barcode
bcs$barcode <- NULL
sobj <- AddMetaData(sobj, metadata = bcs) # add new annotations to meta data slot
```

### Adding Pseudotime to Metadata and save ###

```{r trjectory, fig.height=5, fig.width=6, echo=FALSE} 
# import the pseudo time made with Seurat V5
pseudotime <- readRDS("/Users/hamid/Desktop/ABG MSc project/Raw data/Ileum tissue/Wiarda - Ileum Tissue (non-immune)/pseudotime.Ileum.tissue.rds")
# Add out pseudotime to seurat v4 object that we will later use for co-expression module dynamics with pseudotime
sobj$pseudotime <- pseudotime

# Save the Seurat v4 file
n.cells <- length(sobj@meta.data$seurat_clusters)
orig.cells <- rep("Ileum Tiss", n.cells)
sobj <- AddMetaData(object = sobj, metadata = orig.cells, col.name = 'cell.origin')

# Save the Seurat object
SaveH5Seurat(sobj, overwrite = TRUE, filename = "ileum_pig_tiss(WIERDA).epithel.v4.WGCNA")
```



##### Seurat v4 for WGCNA comparability ####
#### Porcine Ileum Organoid (ABG) ####



#################### Opening the files ####################

# In the chuck below we will make and combine the seurat objects of replicates 
```{r, importing the pre-processed data and merge technical replicates, warning=FALSE, echo=FALSE}
# Define paths
parent_directory = "/Users/hamid/Desktop/ABG MSc project/Raw data/Ileum Org/Non-AGGR"
sample_path = c("ileum_org_pig-F3_IL1-filtered_ref_GTF-incl_intr", "ileum_org_pig-G3_IL1_1-filtered_ref_GTF-incl_intr")
sample_names = c("ileum_Org_1", "ileum_Org_2")
project_name = "Ileum Organoid"

# Load in the pre-processed 10X data
d10x.data <- sapply(sample_path, function(i){
  d10x <- Read10X(file.path(parent_directory, i, "new-outs/counts/FilteredCounts/"))
})

# Make a vector of Seurat Objects and remove cells genes that are not expression in at least 5 cells
sobj_list <- list() 
for (i in 1:length(x = sample_path)) { 
  name <- paste('', sample_path[i], sep='') 
  sobj_list[name] <- list(CreateSeuratObject(counts = d10x.data[[i]], project = paste(sample_names[i]), min.cells = 0, min.features = 0)) 
}

# Merge the Technical replicates (if there is any)
if (length(sample_names) > 1) {
  for (i in 1:length(x = sobj_list)) {
  sobj <- merge(x = sobj_list[[1]], y = sobj_list[[i]]) 
  }
} else {
  sobj <- sobj_list[[1]]
}
rm(sobj_list) # remove un-used list 

# UPDATE GENE NAMES
# Connect to the sScrofa database using mart
ensembl <- useMart("ensembl", dataset="sscrofa_gene_ensembl") 
# For all of our ensembl_gene_id, find all gene symbols
bm <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), values=rownames(sobj), mart=ensembl)
# Find all the ENSEMBL ids in our Seurat object
ENS <- sobj@assays$RNA@counts@Dimnames[[1]][c(grep("ENS", rownames(sobj)))]
# Find the positions of our Seurat ENSEMBL ids in the bm output
ENSEMBL_IDs.pos <- na.omit(c(match(c(paste(ENS)), bm[,1]))) # and remove NAs (ENS with no gene IDs)
# Now find the GeneIDs in bm using the ENSEMBL_IDs.pos
geneIDs <- bm[,2][ENSEMBL_IDs.pos] 
# Since not all ENSEMBL_IDs have true gene name, we need to find the index of geneIDs with no gene name and remove their corresponding ENSEMBL_IDs
# To do this, we will remove genes with no names, and then find the index of these genesID in bm in order to find the corresponding ENSEMBL ids
geneIDs <- as.character(geneIDs[geneIDs !=""])
# Ensure that we don't replace ENSEMBL ids with gene names that could result in duplicates
duplicates <- generics::intersect(geneIDs, rownames(sobj)) # find duplicated between new gene-ids with existing ones in the Seurat object
duplicates.pos <- c(match(c(paste(duplicates)), geneIDs)) # Find postion of the duplicate
geneIDs <- geneIDs[-duplicates.pos] # and remove the duplicarte

geneIDs.pos <- c(match(c(paste(geneIDs)), bm[,2])) # find position of gene IDs in them
ENSEMBL_IDs <- bm[,1][geneIDs.pos] # find the ENSEMBL_IDs using the positions of the gene IDs
# Now we have found all the ENSEMBL ids that have a annotated gene name (winch are also found in our Seurat object)
# We will their final positions in our Seurat object and finally replace them with the true gene ID
# NOTE: The order of the genes are conserved using the method above
final.pos <- c(match(c(paste(ENSEMBL_IDs)), rownames(sobj)))

# replace the name of that genes for all slots
sobj@assays$RNA@counts@Dimnames[[1]][c(final.pos)] <- c(paste(geneIDs))
sobj@assays$RNA@data@Dimnames[[1]][c(final.pos)] <- c(paste(geneIDs))
rownames(sobj@assays$RNA@meta.features)[c(final.pos)] <- c(paste(geneIDs))

# Change the name of the ENSSSCG00000026302 to MKI67
MKI67_gene <- "ENSSSCG00000026302"
# position/index of all genes for all slots
genes <- sobj@assays$RNA@counts@Dimnames[[1]]
# find position of ENSSSCG00000026302
MKI67.position.sobj <- c(match(c(paste(MKI67_gene)), genes))
# replace the name of that gene to MUC2 fro all slots
sobj@assays$RNA@counts@Dimnames[[1]][[MKI67.position.sobj]] <- c(paste("MKI67")) 
sobj@assays$RNA@data@Dimnames[[1]][[MKI67.position.sobj]] <- c(paste("MKI67")) 
rownames(sobj@assays$RNA@meta.features)[MKI67.position.sobj] <- c(paste("MKI67"))

# Change the name of the MT genes to normal (without the "-MT")
MT_genes = c("MT-ND1", "MT-ND2", "MT-COX1", "MT-COX2", "MT-ATP8", "MT-ATP6", "MT-COX3", "MT-ND3", "MT-ND4L", "MT-ND4", "MT-ND5", "MT-ND6", "MT-CYTB")
MT_genes.new = c("ND1", "ND2", "COX1", "COX2", "ATP8", "ATP6", "COX3", "ND3", "ND4L", "ND4", "ND5", "ND6", "CYTB")
 for (i in 1:length(x = MT_genes)) {
   # position/index of all genes for all slots
   genes <- sobj@assays$RNA@counts@Dimnames[[1]]
   # find position of ENSSSCG00000026302
   MT.position.sobj <- c(match(c(paste(MT_genes[i])), genes))
   # replace the name of that gene to the new one for all slots 
   sobj@assays$RNA@counts@Dimnames[[1]][[MT.position.sobj]] <- c(paste(MT_genes.new[i])) 
   sobj@assays$RNA@data@Dimnames[[1]][[MT.position.sobj]] <- c(paste(MT_genes.new[i])) 
   rownames(sobj@assays$RNA@meta.features)[MT.position.sobj] <- c(paste(MT_genes.new[i]))
 }
```

#### Standard Workflow ####

# In the chuck below we will apply the our previous workflow starting with the normalization, feauture selection, sclaing, PCA, clustering and DGE analysis
# For the clustering, we will analyze all clusters manually and select the best.
```{r applying workflow on filtered data, fig.height=4, fig.width=5, echo=FALSE}
# Normalization
sobj <- NormalizeData(sobj, normalization.method = "LogNormalize", scale.factor = 10000) 
# Most variable feature selection
sobj <- FindVariableFeatures(sobj, selection.method = "vst", nfeatures = 2300) # We decreased the number of informative features for organoids as it represent low complexity cells (previously 2500)
# Scale the data
all.genes.sobj <- rownames(sobj)
sobj <- ScaleData(sobj, features = all.genes.sobj) 
# Re-update the new gene names in the scaled data slot after scaling according to the Seruat V3/V4 method
rownames(sobj@assays[["RNA"]]@scale.data)[c(final.pos)] <- c(paste(geneIDs)) 
rownames(sobj@assays[["RNA"]]@scale.data)[MKI67.position.sobj] <- c(paste("MKI67")) 
 for (i in 1:length(x = MT_genes)) {
   # position/index of all genes for all slots
   genes <- sobj@assays$RNA@counts@Dimnames[[1]]
   # find position of ENSSSCG00000026302
   MT.position.sobj <- c(match(c(paste(MT_genes[i])), genes))
   # replace the name of that gene to the new one for all slots 
   rownames(sobj@assays[["RNA"]]@scale.data)[c(MT.position.sobj)] <- c(paste(MT_genes.new[i])) 
 }
# PCA using selected features
sobj <- RunPCA(sobj, features = VariableFeatures(object = sobj)) 
pc.stdev.percentage <- sobj[["pca"]]@stdev / sum(sobj[["pca"]]@stdev) * 100 
cumu.percentage <- cumsum(pc.stdev.percentage) 
co1 <- which(cumu.percentage > 90 & pc.stdev.percentage < 5)[1] 
co2 <- sort(which((pc.stdev.percentage[1:length(pc.stdev.percentage) - 1] - pc.stdev.percentage[2:length(pc.stdev.percentage)]) > 0.01), decreasing = T)[1]  
pcs <- min(co1, co2) 
pc.dims <- 1:pcs
# UMAP/T-SNE clustering
sobj <- FindNeighbors(sobj, dims = pc.dims)
sobj <- FindClusters(sobj, resolution = 1.4) 
sobj <- RunUMAP(sobj, dims = pc.dims) 
```
# In the chuck below, we choose the preleminary cluster based on expectation and visualize
```{r final cluster, fig.height=4, fig.width=5, echo=FALSE} 
# setting final cluster
Idents(sobj) <- "seurat_clusters" # RNA_snn_res.1 or seurat_clusters
DimPlot(sobj, reduction = 'umap', label = TRUE, pt.size = 0.1) + ggtitle(project_name, subtitle = paste("Preleminary Clustering (n = ", paste(dim(sobj)[2], ")", sep = ""))) + ylab("UMAP 1") + xlab("UMAP 2")

sobj@meta.data
```

At this point, we have classified our porcine ileum organoid scRNA-seq dataset and been able to classify clusters into the following cell lineages:

* 'Prolific Stem/Prog'                (cluster 5) 
* 'Prolific TAC'                      (cluster 4) 
* 'TAC'                               (cluster 7, 10)
* 'Early EC Prog'                     (cluster 9)
* 'Late EC Prog'                      (cluster 14)
* 'Resting Cells'                     (cluster 12) 
* 'Undeveloped EC1'                   (cluster 0) 
* 'Undeveloped EC2'                   (cluster 2) 
* 'Renewing EC'                       (cluster 13)
* 'Secretory Lineage'                 (cluster 16)
* 'Developing EC'                     (cluster 3, 8)
* 'Early Immature EC'                 (cluster 6)
* 'Late Immature EC'                  (cluster 1)
* 'Semi-Mature EC'                    (cluster 15)
* 'Migrating EC (Villous Tips)'))     (cluster 11) 

# Preliminary annotation
```{r, annotation plot, fig.height=4, fig.width=7, echo=FALSE}

#### Final annotations ####
bcs <- as.data.frame(colnames(sobj))
colnames(bcs) <- 'barcode'
bcs$cellID <- rep('_', nrow(bcs))
df <- as.data.frame(sobj$seurat_clusters)

# Cluster-based cell annotations not provided in the annotation .txt files:
C5 <- rownames(subset(df, sobj$seurat_clusters %in% c("5")))
C4 <- rownames(subset(df, sobj$seurat_clusters %in% c("4")))
C7.10 <- rownames(subset(df, sobj$seurat_clusters %in% c("7", "10")))
C9 <- rownames(subset(df, sobj$seurat_clusters %in% c("9")))
C14 <- rownames(subset(df, sobj$seurat_clusters %in% c("14")))
C12 <- rownames(subset(df, sobj$seurat_clusters %in% c("12")))
C0.2 <- rownames(subset(df, sobj$seurat_clusters %in% c("0", "2")))
C13 <- rownames(subset(df, sobj$seurat_clusters %in% c("13")))
C16 <- rownames(subset(df, sobj$seurat_clusters %in% c("16")))
C3.8 <- rownames(subset(df, sobj$seurat_clusters %in% c("3", "8")))
C6 <- rownames(subset(df, sobj$seurat_clusters %in% c("6")))
C1 <- rownames(subset(df, sobj$seurat_clusters %in% c("1")))
C15 <- rownames(subset(df, sobj$seurat_clusters %in% c("15")))
C11 <- rownames(subset(df, sobj$seurat_clusters %in% c("11")))

# name the cells clusters (rename for ML)
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C5, 'Prolific Stem/Prog')) 
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C4, 'Prolific TAC'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C7.10, 'TAC'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C9, 'Early EC Prog'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C14, 'Late EC Prog')) 
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C12, 'Resting Cells'))
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C0.2, 'Undeveloped EC')) 
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C13, 'Renewing EC')) 
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C16, 'Secretory Lineage')) 
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C3.8, 'Developing EC')) 
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C6, 'Early Immature EC')) 
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C1, 'Late Immature EC')) 
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C15, 'Semi-Mature EC')) 
bcs <- bcs %>% mutate(cellID = replace(cellID, barcode %in% C11, 'Migrating EC (Villous Tips)'))

# update the metadatas slot
rownames(bcs) <- bcs$barcode
bcs$barcode <- NULL
sobj <- AddMetaData(sobj, metadata = bcs) # add new annotations to meta data slot

#######   Update secretory lineage   #############

# Take Paneth cells
pan1 <- WhichCells(sobj[,sobj$cellID == "Secretory Lineage"], expression = RETNLB > 0 & FCN2 > 0)
pan2 <- WhichCells(sobj[,sobj$cellID == "Secretory Lineage"], expression = SOX9 > 0)
pan.combined <- c(pan1, pan2) # combine
pan.unique <- unique(pan.combined) # and keep remove duplicates
# Take Goblet cells that are NOT Paneth (since MUC2 appears not to be Goblet specific in Pig (see literature), we consider MUC2+ cells lacking Paneth signature as Goblet
gob1 <- WhichCells(sobj[,sobj$cellID == "Secretory Lineage"], expression = MUC2 > 0 & TFF3 > 0)
gob.unique <-  gob1[is.na(match(gob1,pan.unique))] # Find cells that are not in Paneth
# Take Enteroendocrine
EEC.unique <- WhichCells(sobj[,sobj$cellID == "Secretory Lineage"], expression = CHGA  > 0 | NEUROD1 > 0  | NEUROG3 > 0  | PAX4 > 0)
# Take secretory progenitor that are not in PC, BC and EEC
prog <- WhichCells(sobj[,sobj$cellID == "Secretory Lineage"], expression = MKI67 | TK1 | TOP2A | PLK1 | CCNB1 | HMGB2 | TPX2 | H2AC6 | H2AC4 | H2AC20 | CENPF | CDC20 | NUSAP1 | UBE2C > 0)
prog.unique.in.pan <- prog[is.na(match(prog, pan.unique))] # keep ones that are not in paneth
prog.unique.in.gob <- prog.unique.in.pan[is.na(match(prog.unique.in.pan, gob.unique))] # keep ones that are not in goblet
prog.unique.in.EEC <- prog.unique.in.gob[is.na(match(prog.unique.in.gob, EEC.unique))] # keep ones that are not in EEC
prog.unique <- as.vector(na.omit(prog.unique.in.EEC)) # keep unique and remove NA

# Adjusting the annotation of some secretory cells in the Ident slot
Idents(sobj, cells = pan.unique) <- "PC"
Idents(sobj, cells = gob.unique) <- "GC" 
Idents(sobj, cells = EEC.unique) <- "EEC"
Idents(sobj, cells = prog.unique) <- "PC" # change to PC in case it's a paneth

# adjusting the names in the remaining slots (metadata and sobj$cellid)
# Finding the positions of the cell types
PC.pos <- c(match(c(paste(pan.unique)), rownames(sobj@meta.data)))
GC.pos <- c(match(c(paste(gob.unique)), rownames(sobj@meta.data)))
EEC.pos <- c(match(c(paste(EEC.unique)), rownames(sobj@meta.data)))
prog.pos <- c(match(c(paste(prog.unique)), rownames(sobj@meta.data)))

# Find position/index of all cell in the "secretory cluster"
secr.cells.pos <- which(sobj@meta.data[["cellID"]] == "Secretory Lineage")
secr.cells <- sobj@meta.data[["cellID"]][c(secr.cells.pos)]

# replace the name of that gene celltype
sobj@meta.data[["cellID"]][PC.pos] <- c(paste("PC")) 
sobj@meta.data[["cellID"]][GC.pos] <- c(paste("GC")) 
sobj@meta.data[["cellID"]][EEC.pos] <- c(paste("EEC")) 
sobj@meta.data[["cellID"]][prog.pos] <- c(paste("PC")) # change to PC in case it's a paneth
# change remaining cell (far from secretory lineage cluster) to goblet
sobj@meta.data[["cellID"]][which(sobj@meta.data[["cellID"]] == "Secretory Lineage")] <- c(paste("GC")) # its position 480
```

### Adding Pseudotime to Metadata and save ###

```{r trjectory, fig.height=5, fig.width=6, echo=FALSE} 
# import the organoid pseudo time made with Seurat V5
pseudotime <- readRDS("/Users/hamid/Desktop/ABG MSc project/Raw data/Ileum Org/Decon Experimental/pseudotime.Ileum.Organoid.rds")
# Add out pseudotime to seurat v4 object that we will later use for co-expression module dynamics with pseudotime
sobj$pseudotime <- pseudotime

# Save the Seurat v4 file
n.cells <- length(sobj@meta.data$seurat_clusters)
orig.cells <- rep("Ileum Org", n.cells)
sobj <- AddMetaData(object = sobj, metadata = orig.cells, col.name = 'cell.origin')

# Save the Seurat object
SaveH5Seurat(sobj, overwrite = TRUE, filename = "ileum_pig_org.v4.WGCNA")
```